---
title: "demarray"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demmap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(demarray)
```

## Introduction

## op

`x` is any array with complete metadata

| Condition                    | `y` is Counts | `y` is Values | `y` is array |
|:-----------------------------|:--------------|:-------- -----|:-------------|
| `x` has dimensions not present in `y` | Throw errror | Add dimensions to `y` | Throw error |
| `y` has dimensions not present in `x` | Collapse dimensions in `y` | Throw | Throw error |
| On shared dimension, categories for `x` have many-to-one relationship to categories for `y` | Throw error | Split categories in `y` to match `x` | Throw error |
| On shared dimension, categories for `y` have many-to-one relationship to categories for `y` | Collapse categories in `y` to match `x` | Throw error | Throw error |




# Labels

## Default

| dimlabels       |  example             |
|:----------------|:---------------------|
| Categories      | "New Zealand", "Australia" |

- only restriction is that unique and non-blank (NA allowed)
- iterations are categories, convention that use 1, 2, 3


## Integers

"2020", "2021", "2022"
"<0", "5", "10", "15+"
Cannot have duplicates. Can have open, NA

## Quantities

Plain integers
"2020", "5", "0"

Groups
"1-4", "5-9", "100-999"

Open
"<2000", "100+"

Mixtures
"0", "1-4", "5-9", "10+"

Not necessarily covering whole range
"0-4", "5-9", "15-19", "20-24"


## Intervals

Intervals *deliberately* ambiguous more ambiguous than DateRanges. Conventional labels for years are ambiguous. Using intervals, can represent info in labels without forcing user to clarify, except when essential to do so.

"<2000", "2000-2005", "2005-2010", "2010-2015", "2015-2010", 2020+", "2000-2001"


## Calendar

| dimlabel        |  example             |
|:----------------|:---------------------|
| Quarters         | "<1990 Q4", 2000 Q1", "2020 Q3+" |
| Months          | "<2000 Jan", "2020 Aug", "2025 Feb+" |
| DatePoints          | "2020-01-01", "2005-06-01" |
| DateRanges         | "[2020-01-01, 2020-12-31]", (-Inf, 2019-12-31]" |



## Specialised

| dimlabels     | dimtype   |  example      |
|:--------------|:----------|:--------------|
| Triangles     | triangles | "Lower", "Upper" |
| Directions    | directions | "In", "Out"      |
| Quantiles     | quantiles | "2.5%", "50%", "97.5%" |

- if 'strict' is TRUE, dimensions with the indicated dimtype must have the associated dimlabels
- can use these dimlabels with other dimtypes


## Notes

Messy
- aim is to represent messy labeling conventions - inherits messiness from the conventions
- tangled relationship
  - everything a special case of Categories
- original labeling conventions ambiguous. Deliberately carry that ambiguity over to the labels
- "lazy disambiguation" - only disambiguate when we need to


rather than impose new unambiguous labelling system, allow people to work with existing ambiguous system